---
output: 
  html_document
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(gradethis)
knitr::opts_chunk$set(echo = FALSE)
```

```{r include=FALSE}
library(limma)
```

```{r}
load("../data/data.Rdata")
```

**Challenge**

1. Add another contrast to the contrasts matrix: `L.PregVsLac = luminal.pregnant - luminal.lactate` and re-run the code above. You should have two comparisons in `fit.cont` now.
2. Check out the `vennDiagram` function
(HINT: type `?vennDiagram`).
Can you show the overlap of differentially expressed genes between the two comparisons? How many genes are commonly differentially expressed?

```{r challenge5, exercise=TRUE}
cont.matrix <- makeContrasts(B.PregVsLac=____,
                             L.PregVsLac=____,
                             levels=____)
fit.cont <- ____
fit.cont <- ____
summa.fit <- ____
summary(____)
# Venn diagram
par(mfrow=c(1,1))
vennDiagram(____,include=____,counts.col=c("purple", "black"),
    circle.col = c("blue", "green3"))
```

```{r challenge5-solution, exercise.reveal_solution = FALSE}
cont.matrix <- makeContrasts(B.PregVsLac=basal.pregnant - basal.lactate,
                             L.PregVsLac=luminal.pregnant - luminal.lactate,
                             levels=design)
fit.cont <- contrasts.fit(fit, cont.matrix)
fit.cont <- eBayes(fit.cont)
summa.fit <- decideTests(fit.cont)
summary(summa.fit)
# Venn diagram
par(mfrow=c(1,1))
vennDiagram(summa.fit,include=c("up", "down"),counts.col=c("purple", "black"),
    circle.col = c("blue", "green3"))
```

```{r challenge5-code-check}
grade_this_code()
```