---
output: 
  html_document
runtime: shiny_prerendered
---

```{r setup12, include=FALSE}
library(learnr)
library(gradethis)
knitr::opts_chunk$set(echo = FALSE)
```

```{r include=FALSE}
library(limma)
```

```{r}
if(!exists("sampleinfo")) {
    load("../data/data.Rdata")
}
```

### Challenge 12

1. Produce a barcodeplot for luminal pregnant vs lactating for the gene set DANG_REGULATED_BY_MYC_UP. Does the pattern of enrichment look as strong?

```{r challenge12_1, exercise=TRUE}
barcodeplot(____, main="L.PregVsLac: DANG_REGULATED_BY_MYC_UP")
````

```{r challenge12_1-solution, exercise.reveal_solution = FALSE}
barcodeplot(fit.cont$coeff[,2], index=c2.ind[["DANG_REGULATED_BY_MYC_UP"]], main="L.PregVsLac: DANG_REGULATED_BY_MYC_UP")
```

```{r challenge12_1-code-check}
grade_this_code()
```

2. Produce a barcode plot for the Wnt signalling pathway REACTOME_REPRESSION_OF_WNT_TARGET_GENES

```{r challenge12_2, exercise=TRUE}
barcodeplot(____, main="L.PregVsLac: REACTOME_REPRESSION_OF_WNT_TARGET_GENES")
```

```{r challenge12_2-solution, exercise.reveal_solution = FALSE}
barcodeplot(fit.cont$coeff[,2], index=c2.ind[["REACTOME_REPRESSION_OF_WNT_TARGET_GENES"]], main="L.PregVsLac: REACTOME_REPRESSION_OF_WNT_TARGET_GENES")
```

```{r challenge12_2-code-check}
grade_this_code()
```

3. You can put two gene sets on one plot, for example a set that is up-regulated and one that is down-regulated,
by adding a gene set to the `index2` argument. Produce a barcodeplot with the two sets REACTOME_REPRESSION_OF_WNT_TARGET_GENES and PID_WNT_CANONICAL_PATHWAY.

```{r challenge12_3, exercise=TRUE}
barcodeplot(____, main="L.PregVsLac: REACTOME_REPRESSION_OF_WNT_TARGET_GENES and PID_WNT_CANONICAL_PATHWAY sets")
````

```{r challenge12_3-solution, exercise.reveal_solution = FALSE}
barcodeplot(fit.cont$coeff[,2], index=c2.ind[["REACTOME_REPRESSION_OF_WNT_TARGET_GENES"]], index2=c2.ind[["PID_WNT_CANONICAL_PATHWAY"]], main="L.PregVsLac: REACTOME_REPRESSION_OF_WNT_TARGET_GENES and PID_WNT_CANONICAL_PATHWAY sets")
```

```{r challenge12_3-code-check}
grade_this_code()
```

<br>

